### 2022-05-19 README.Rmd
encoding woes

The approach uses quasi-random sampling (Grafston & Tille, 2013, Foster et al., 2018) to generate a quadrature scheme for numerical approximation of a Poisson point process model (Berman & Turner 1992; Warton & Shepard 2010). Quasi-random sampling quadrature are form of spatially-balanced survey design or point stratification that aims to reduce the frequency of placing samples close to each other (relative to pseudo-random or grid designs). A quasi-random quadrature design improves efficiency of background point sampling (and subsequent modelling) by reducing the amount of spatial auto-correlation between data implying that each sample is providing as much unique information as possible (Grafston & Tille, 2013, Foster et al., 2018) and thus reducing low errors for geostatistical prediction (Diggle & Ribeiro, 2007). Because the quasi-random design is not on a regular grid we use a Dirichlet tessellation to generate polygons for each point. Areal weights are then derived from these polygons. Under this approach we **do not** use a targeted background scheme for generating 'pseudo-absences', but rather can account for bias in the sighting process
of presences using covariates or a offset when modelling (e.g Warton et al., 2013; Fithian et al, 2015).

@@

Diggle, P. J., P. J. Ribeiro, Model-based Geostatistics. Springer Series in Statistics. Springer, 2007.

Foster, S.D., Monk, J., Lawrence, E., Hayes, K.R., Hosack, G.R. and Przeslawski, R., 2018. Statistical Considerations for Monitoring and Sampling.

Grafstrom, Anton, and Yves Tille. "Doubly balanced spatial sampling with spreading and restitution of auxiliary totals." Environmetrics 24.2 (2013): 120-131.

Renner, I.W., Elith, J., Baddeley, A., Fithian, W., Hastie, T., Phillips, S.J., Popovic, G. and Warton, D.I., 2015. Point process models for presence only analysis. Methods in Ecology and Evolution, 6(4), p.366-379.

Warton, D. I., and L. C. Shepherd. "Poisson point process models solve the pseudo-absence problem for presence-only data in ecology." The Annals of Applied Statistics 4.3 (2010): 1383-1402.

Warton, D.I., Renner, I.W. and Ramp, D., 2013. Model-based control of observer bias for the analysis of presence-only data in ecology. PloS one, 8(11), p.e79168.

>>>>

The approach uses quasi-random sampling (Grafston & Tille, 2013, Foster et al., 2018) to generate a quadrature scheme for numerical approximation of a Poisson point process model (Berman & Turner 1992; Warton & Shepard 2010). Quasi-random sampling quadrature are form of spatially-balanced survey design or point stratification that aims to reduce the frequency of placing samples close to each other (relative to pseudo-random or grid designs). A quasi-random quadrature design improves efficiency of background point sampling (and subsequent modelling) by reducing the amount of spatial auto-correlation between data implying that each sample is providing as much unique information as possible (Grafston & Tille, 2013, Foster et al., 2018) and thus reducing low errors for geostatistical prediction (Diggle & Ribeiro, 2007). Because the quasi-random design is not on a regular grid we use Dirichlet tessellation to generate polygons for each point in the quadrature scheme. Areal weights are then derived from these polygons.

We note that this approach is different to a targeted background scheme for generating pseudo-absences. If the users intent is to reduce sighting biases via a targeted background scheme, we recommend that bias
is accounted for via covariates (distance to roads) or an offset (a known amount of effort) (e.g Warton et al., 2013;Fithian et al, 2015).

@@

Diggle, P. J., P. J. Ribeiro, *Model-based Geostatistics*. Springer Series in Statistics. Springer, 2007.

Foster, S.D., Monk, J., Lawrence, E., Hayes, K.R., Hosack, G.R. and Przeslawski, R., 2018. *Statistical considerations for monitoring and sampling.* Field manuals for marine sampling to monitor Australian waters, pp.23-41.

Grafstrom, Anton, and Yves Tille. *Doubly balanced spatial sampling with spreading and restitution of auxiliary totals.* Environmetrics 24.2 (2013): 120-131.

Renner, I.W., Elith, J., Baddeley, A., Fithian, W., Hastie, T., Phillips, S.J., Popovic, G. and Warton, D.I., 2015. *Point process models for presence only analysis.* Methods in Ecology and Evolution, 6(4), p.366-379.

Warton, D. I., and L. C. Shepherd. *Poisson point process models solve the pseudo-absence problem for presence-only data in ecology.* The Annals of Applied Statistics 4.3 (2010): 1383-1402.

Warton, D.I., Renner, I.W. and Ramp, D., 2013. *Model-based control of observer bias for the analysis of presence-only data in ecology.* PloS one, 8(11), p.e79168.
